---
title: "Compute per-barcode binding to mAb samples"
author: "Tyler Starr"
date: "5/12/2022"
output:
  github_document:
    html_preview: false
editor_options: 
  chunk_output_type: inline
---

This notebook reads in per-barcode counts from `count_variants.ipynb` for sera-binding titration experiments, computes functional scores for RBD binding values via delta-AUC metrics, and does some basic QC on variant binding functional scores.

```{r setup, message=FALSE, warning=FALSE, error=FALSE}
#list of packages to install/load
packages = c("yaml","data.table","tidyverse","gridExtra")
#install any packages not already installed
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == F)){
  install.packages(packages[!installed_packages],
                   lib=c(paste("/uufs/chpc.utah.edu/common/home/",Sys.getenv("USER"),"/RLibs/",Sys.getenv("R_VERSION"),sep="")),
                   repos=c("http://cran.us.r-project.org"))
}
#load packages
invisible(lapply(packages, library, character.only=T))

knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(dev.args = list(png = list(type = "cairo")))

#read in config file
config <- read_yaml("config.yaml")

#make output directory
if(!file.exists(config$mAb_EC50_dir)){
  dir.create(file.path(config$mAb_EC50_dir))
}
```
Session info for reproducing environment:
```{r print_sessionInfo}
sessionInfo()
```

## Setup
First, we will read in metadata on our sort samples, the table giving number of reads of each barcode in each of the sort bins, and the barcode-variant lookup tables, and merge these tables together.

```{r input_data}
#read dataframe with list of barcode runs
barcode_runs <- read.csv(file=config$barcode_runs,stringsAsFactors=F); barcode_runs <- subset(barcode_runs, select=-c(R1))

#read file giving count of each barcode in each sort partition
counts <- data.table(read.csv(file=config$variant_counts_file,stringsAsFactors=F))

#read in barcode-variant lookup tables
dt <- data.table(read.csv(file=config$codon_variant_table_file_lib61,stringsAsFactors=F))
setkey(dt,barcode,library)

dt <- merge(counts, dt, by=c("library","barcode")); rm(counts)

samples_C68_61 <- data.frame(sample=sort(unique(paste(rep("C68_61",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_61","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_183 <- data.frame(sample=sort(unique(paste(rep("C68_183",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_183","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_185 <- data.frame(sample=sort(unique(paste(rep("C68_185",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_185","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_3 <- data.frame(sample=sort(unique(paste(rep("C68_3",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_61","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_10 <- data.frame(sample=sort(unique(paste(rep("C68_10",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_10","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_88 <- data.frame(sample=sort(unique(paste(rep("C68_88",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_88","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_201 <- data.frame(sample=sort(unique(paste(rep("C68_201",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_201","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_239 <- data.frame(sample=sort(unique(paste(rep("C68_239",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_239","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

samples_C68_490 <- data.frame(sample=sort(unique(paste(rep("C68_490",6),formatC(barcode_runs[barcode_runs$sample_type=="C68_490","concentration"], width=2,flag="0"),sep="_"))),conc=c(10000, 400, 16, 0.64, 0.0256, 0))

```

 Convert from Illumina read counts to estimates of the number of cells that were sorted into a bin, and add some other useful information to our data frame.
 
```{r downweight_counts_by_cells}
#for each bin, normalize the read counts to the observed ratio of cell recovery among bins
for(i in 1:nrow(barcode_runs)){
  lib <- as.character(barcode_runs$library[i])
  bin <- as.character(barcode_runs$sample[i])
  ratio <- sum(dt[library==lib & sample==bin,"count"])/barcode_runs$number_cells[i]
  if(ratio<1){ #if there are fewer reads from a FACS bin than cells sorted
    dt[library==lib & sample==bin, count.norm := as.numeric(count)] #don't normalize cell counts, make count.norm the same as count
    print(paste("reads < cells for",lib,bin,", un-normalized (ratio",ratio,")")) #print to console to inform of undersampled bins
  }else{
    dt[library==lib & sample==bin, count.norm := as.numeric(count/ratio)] #normalize read counts by the average read:cell ratio, report in new "count.norm" column
    print(paste("read:cell ratio for",lib,bin,"is",ratio))
  }
}

#annotate each barcode as to whether it's a homolog variant, wildtype, synonymous muts only, stop, nonsynonymous, >1 nonsynonymous mutations
dt[,variant_class:=as.character(NA)]
dt[n_codon_substitutions==0, variant_class := "wildtype"]
dt[n_codon_substitutions > 0 & n_aa_substitutions==0, variant_class := "synonymous"]
dt[n_aa_substitutions>0 & grepl("*",aa_substitutions,fixed=T), variant_class := "stop"]
dt[n_aa_substitutions == 1 & !grepl("*",aa_substitutions,fixed=T), variant_class := "1 nonsynonymous"]
dt[n_aa_substitutions > 1 & !grepl("*",aa_substitutions,fixed=T), variant_class := ">1 nonsynonymous"]

#cast the data frame into wide format
dt <- dcast(dt, library + barcode + target + variant_class + aa_substitutions + n_aa_substitutions ~ sample, value.var="count.norm")

```

## Calculating mean bin for each barcode at each sample concentration
Next, for each barcode at each of the sera concentrations, calculate the "mean bin" response variable. This is calculated as a simple mean, where the value of each bin is the integer value of the bin (bin1=unbound, bin4=highly bound) -- because of how bins are defined, the mean fluorescence of cells in each bin are equally spaced on a log-normal scale, so mean bin correlates with simple mean fluorescence.

We do not use the fluorescence boundaries of the FACS bins in our calculations here, but we provide them for posterity's sake below.

`(-288, 53), (53, 1099), (1099, 22649), (22649, 262143)`

```{r calculate_mean_bin}
#function that returns mean bin and sum of counts for four bins cell counts. Includes cutoffs for bimodal sample splits to filter out
calc.meanbin <- function(vec, split13filter=0.4, split24filter=0.4, split14filter=0.2){
  total <- sum(vec)
  if(is.na(total) | (vec[1] > split13filter*total & vec[3] > split13filter*total) | (vec[2] > split24filter*total & vec[4] > split24filter*total) | (vec[1] > split14filter*total & vec[4] > split14filter*total)){
    return(list(NA,NA))
  }else{
    return( list((vec[1]*1+vec[2]*2+vec[3]*3+vec[4]*4)/(vec[1]+vec[2]+vec[3]+vec[4]), total) )
  }
}
  

#iterate through titration samples, compute mean_bin and total_count for each barcode variant
#C68_61
for(i in 1:nrow(samples_C68_61)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_61[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_61[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_61[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_61[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_61[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_61[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_183
for(i in 1:nrow(samples_C68_183)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_183[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_183[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_183[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_183[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_183[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_183[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_185
for(i in 1:nrow(samples_C68_185)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_185[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_185[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_185[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_185[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_185[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_185[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_3
for(i in 1:nrow(samples_C68_3)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_3[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_3[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_3[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_3[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_3[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_3[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_10
for(i in 1:nrow(samples_C68_10)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_10[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_10[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_10[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_10[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_10[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_10[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_88
for(i in 1:nrow(samples_C68_88)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_88[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_88[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_88[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_88[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_88[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_88[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_201
for(i in 1:nrow(samples_C68_201)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_201[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_201[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_201[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_201[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_201[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_201[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_239
for(i in 1:nrow(samples_C68_239)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_239[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_239[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_239[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_239[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_239[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_239[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}

#C68_490
for(i in 1:nrow(samples_C68_490)){ #iterate through titeseq sample (concentration)
  meanbin_out <- paste(samples_C68_490[i,"sample"],"_meanbin",sep="") #define the header name for the meanbin output for the given concentration sample
  totalcount_out <- paste(samples_C68_490[i,"sample"],"_totalcount",sep="") #define the header name for the total cell count output for the given concentration sample
  bin1_in <- paste(samples_C68_490[i,"sample"],"_bin1",sep="") #define the header names for the input cell counts for bins1-4 of the given concnetration sample
  bin2_in <- paste(samples_C68_490[i,"sample"],"_bin2",sep="")
  bin3_in <- paste(samples_C68_490[i,"sample"],"_bin3",sep="")
  bin4_in <- paste(samples_C68_490[i,"sample"],"_bin4",sep="")
  dt[,c(meanbin_out,totalcount_out) := calc.meanbin(c(get(bin1_in),get(bin2_in),get(bin3_in),get(bin4_in))),by=c("barcode","library")]
}
```

## Fit binding curves
We will calculate a simple AUC metric across each barcode's titration series. We will also include a minimum cell count that is required for a meanbin estimate to be used in the titration fit, and a minimum number of concentrations with determined meanbin that is required for a titration to be reported. 

```{r fit_titrations}
#For QC and filtering, output columns giving the average number of cells that were sampled for a barcode across the 9 sample concentrations, and a value for the number of meanbin estimates that were removed for being below the # of cells cutoff
cutoff <- 2

dt[,`C68_61_avgcount` := mean(c(`C68_61_01_totalcount`,`C68_61_02_totalcount`,`C68_61_03_totalcount`,
                             `C68_61_04_totalcount`,`C68_61_05_totalcount`,`C68_61_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_61_min_cell_filtered` := sum(c(c(`C68_61_01_totalcount`,`C68_61_02_totalcount`,`C68_61_03_totalcount`,
                                       `C68_61_04_totalcount`,`C68_61_05_totalcount`,`C68_61_06_totalcount`)<cutoff,
                                     is.na(c(`C68_61_01_totalcount`,`C68_61_02_totalcount`,`C68_61_03_totalcount`,
                                             `C68_61_04_totalcount`,`C68_61_05_totalcount`,`C68_61_06_totalcount`))),na.rm=T),by=c("library","barcode")]

dt[,`C68_183_avgcount` := mean(c(`C68_183_01_totalcount`,`C68_183_02_totalcount`,`C68_183_03_totalcount`,
                             `C68_183_04_totalcount`,`C68_183_05_totalcount`,`C68_183_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_183_min_cell_filtered` := sum(c(c(`C68_183_01_totalcount`,`C68_183_02_totalcount`,`C68_183_03_totalcount`,
                                       `C68_183_04_totalcount`,`C68_183_05_totalcount`,`C68_183_06_totalcount`)<cutoff,
                                     is.na(c(`C68_183_01_totalcount`,`C68_183_02_totalcount`,`C68_183_03_totalcount`,
                                             `C68_183_04_totalcount`,`C68_183_05_totalcount`,`C68_183_06_totalcount`))),na.rm=T),by=c("library","barcode")]

dt[,`C68_185_avgcount` := mean(c(`C68_185_01_totalcount`,`C68_185_02_totalcount`,`C68_185_03_totalcount`,
                             `C68_185_04_totalcount`,`C68_185_05_totalcount`,`C68_185_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_185_min_cell_filtered` := sum(c(c(`C68_185_01_totalcount`,`C68_185_02_totalcount`,`C68_185_03_totalcount`,
                                       `C68_185_04_totalcount`,`C68_185_05_totalcount`,`C68_185_06_totalcount`)<cutoff,
                                     is.na(c(`C68_185_01_totalcount`,`C68_185_02_totalcount`,`C68_185_03_totalcount`,
                                             `C68_185_04_totalcount`,`C68_185_05_totalcount`,`C68_185_06_totalcount`))),na.rm=T),by=c("library","barcode")]

dt[,`C68_3_avgcount` := mean(c(`C68_3_01_totalcount`,`C68_3_02_totalcount`,`C68_3_03_totalcount`,
                             `C68_3_04_totalcount`,`C68_3_05_totalcount`,`C68_3_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_3_min_cell_filtered` := sum(c(c(`C68_3_01_totalcount`,`C68_3_02_totalcount`,`C68_3_03_totalcount`,
                                       `C68_3_04_totalcount`,`C68_3_05_totalcount`,`C68_3_06_totalcount`)<cutoff,
                                     is.na(c(`C68_3_01_totalcount`,`C68_3_02_totalcount`,`C68_3_03_totalcount`,
                                             `C68_3_04_totalcount`,`C68_3_05_totalcount`,`C68_3_06_totalcount`))),na.rm=T),by=c("library","barcode")]



dt[,`C68_10_avgcount` := mean(c(`C68_10_01_totalcount`,`C68_10_02_totalcount`,`C68_10_03_totalcount`,
                             `C68_10_04_totalcount`,`C68_10_05_totalcount`,`C68_10_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_10_min_cell_filtered` := sum(c(c(`C68_10_01_totalcount`,`C68_10_02_totalcount`,`C68_10_03_totalcount`,
                                       `C68_10_04_totalcount`,`C68_10_05_totalcount`,`C68_10_06_totalcount`)<cutoff,
                                     is.na(c(`C68_10_01_totalcount`,`C68_10_02_totalcount`,`C68_10_03_totalcount`,
                                             `C68_10_04_totalcount`,`C68_10_05_totalcount`,`C68_10_06_totalcount`))),na.rm=T),by=c("library","barcode")]


dt[,`C68_88_avgcount` := mean(c(`C68_88_01_totalcount`,`C68_88_02_totalcount`,`C68_88_03_totalcount`,
                             `C68_88_04_totalcount`,`C68_88_05_totalcount`,`C68_88_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_88_min_cell_filtered` := sum(c(c(`C68_88_01_totalcount`,`C68_88_02_totalcount`,`C68_88_03_totalcount`,
                                       `C68_88_04_totalcount`,`C68_88_05_totalcount`,`C68_88_06_totalcount`)<cutoff,
                                     is.na(c(`C68_88_01_totalcount`,`C68_88_02_totalcount`,`C68_88_03_totalcount`,
                                             `C68_88_04_totalcount`,`C68_88_05_totalcount`,`C68_88_06_totalcount`))),na.rm=T),by=c("library","barcode")]


dt[,`C68_201_avgcount` := mean(c(`C68_201_01_totalcount`,`C68_201_02_totalcount`,`C68_201_03_totalcount`,
                             `C68_201_04_totalcount`,`C68_201_05_totalcount`,`C68_201_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_201_min_cell_filtered` := sum(c(c(`C68_201_01_totalcount`,`C68_201_02_totalcount`,`C68_201_03_totalcount`,
                                       `C68_201_04_totalcount`,`C68_201_05_totalcount`,`C68_201_06_totalcount`)<cutoff,
                                     is.na(c(`C68_201_01_totalcount`,`C68_201_02_totalcount`,`C68_201_03_totalcount`,
                                             `C68_201_04_totalcount`,`C68_201_05_totalcount`,`C68_201_06_totalcount`))),na.rm=T),by=c("library","barcode")]


dt[,`C68_239_avgcount` := mean(c(`C68_239_01_totalcount`,`C68_239_02_totalcount`,`C68_239_03_totalcount`,
                             `C68_239_04_totalcount`,`C68_239_05_totalcount`,`C68_239_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_239_min_cell_filtered` := sum(c(c(`C68_239_01_totalcount`,`C68_239_02_totalcount`,`C68_239_03_totalcount`,
                                       `C68_239_04_totalcount`,`C68_239_05_totalcount`,`C68_239_06_totalcount`)<cutoff,
                                     is.na(c(`C68_239_01_totalcount`,`C68_239_02_totalcount`,`C68_239_03_totalcount`,
                                             `C68_239_04_totalcount`,`C68_239_05_totalcount`,`C68_239_06_totalcount`))),na.rm=T),by=c("library","barcode")]


dt[,`C68_490_avgcount` := mean(c(`C68_490_01_totalcount`,`C68_490_02_totalcount`,`C68_490_03_totalcount`,
                             `C68_490_04_totalcount`,`C68_490_05_totalcount`,`C68_490_06_totalcount`),na.rm=T),by=c("library","barcode")]

#number of concentrations at which meanbin is calculated from < cutoff cells or is missing b/c filtered for bimodality
dt[,`C68_490_min_cell_filtered` := sum(c(c(`C68_490_01_totalcount`,`C68_490_02_totalcount`,`C68_490_03_totalcount`,
                                       `C68_490_04_totalcount`,`C68_490_05_totalcount`,`C68_490_06_totalcount`)<cutoff,
                                     is.na(c(`C68_490_01_totalcount`,`C68_490_02_totalcount`,`C68_490_03_totalcount`,
                                             `C68_490_04_totalcount`,`C68_490_05_totalcount`,`C68_490_06_totalcount`))),na.rm=T),by=c("library","barcode")]

#function that fits a nls regression to the titration series, including an option to filter below certain thresholds for average cells across all samples, and number of samples below a cutoff of cells
fit.titration <- function(y.vals,x.vals,count.vals,min.cfu=cutoff,
                          min.means=1,min.average=7,EC50.start=10,
                          a.start=3,a.lower=2,a.upper=3,
                          b.start=1,b.lower=1,b.upper=1.5,
                          n.start=1,n.lower=1/2,n.upper=2){
  indices <- count.vals>min.cfu & !is.na(y.vals)
  y <- y.vals[indices]
  x <- x.vals[indices]
  if((length(y) < min.means*length(y.vals)) | (mean(count.vals,na.rm=T) < min.average)){ #return NAs if < min.means fraction of concentrations have above min.cfu counts or if the average count across all concentrations is below min.average
    return(list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA)))
  }else{
    fit <- nls(y ~ (a/(1+(EC50/x)^n))+b,
               start=list(a=a.start,b=b.start,n=n.start,EC50=EC50.start),
               lower=list(a=a.lower,b=b.lower,n=n.lower,EC50=min(x.vals[x.vals>0])/1000), #constrain EC50 to be no lower than 1/100x the lowest concentration value
               upper=list(a=a.upper,b=b.upper,n=n.upper,EC50=max(x.vals[x.vals>0])*10), #constrain EC50 to be no higher than the 10x highest concentration value
               algorithm="port")
    y.pred <- predict(fit,newdata=list(x=x))
    resid <- y - y.pred
    resid.norm <- resid/as.numeric(summary(fit)$coefficients["a","Estimate"])
    nMSR <- mean((resid.norm)^2,na.rm=T)
    return(list(as.numeric(summary(fit)$coefficients["EC50","Estimate"]),
                as.numeric(summary(fit)$coefficients["EC50","Std. Error"]),
                as.numeric(summary(fit)$coefficients["a","Estimate"]),
                as.numeric(summary(fit)$coefficients["b","Estimate"]),
                as.numeric(nMSR)))
  }
}

#fit titration to C68_61 data for each barcode
dt[,c("EC50_C68_61","EC50_SE_C68_61","response_C68_61","baseline_C68_61","nMSR_C68_61") :=
     tryCatch(fit.titration(y.vals=c(`C68_61_01_meanbin`,`C68_61_02_meanbin`,`C68_61_03_meanbin`,
                                     `C68_61_04_meanbin`,`C68_61_05_meanbin`,`C68_61_06_meanbin`),
                            x.vals=samples_C68_61$conc,
                            count.vals=c(`C68_61_01_totalcount`,`C68_61_02_totalcount`,`C68_61_03_totalcount`,
                                         `C68_61_04_totalcount`,`C68_61_05_totalcount`,`C68_61_06_totalcount`)),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_183 data for each barcode
dt[,c("EC50_C68_183","EC50_SE_C68_183","response_C68_183","baseline_C68_183","nMSR_C68_183") :=
     tryCatch(fit.titration(y.vals=c(`C68_183_01_meanbin`,`C68_183_02_meanbin`,`C68_183_03_meanbin`,
                                     `C68_183_04_meanbin`,`C68_183_05_meanbin`,`C68_183_06_meanbin`),
                            x.vals=samples_C68_183$conc,
                            count.vals=c(`C68_183_01_totalcount`,`C68_183_02_totalcount`,`C68_183_03_totalcount`,
                                         `C68_183_04_totalcount`,`C68_183_05_totalcount`,`C68_183_06_totalcount`)),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_185 data for each barcode
dt[,c("EC50_C68_185","EC50_SE_C68_185","response_C68_185","baseline_C68_185","nMSR_C68_185") :=
     tryCatch(fit.titration(y.vals=c(`C68_185_01_meanbin`,`C68_185_02_meanbin`,`C68_185_03_meanbin`,
                                     `C68_185_04_meanbin`,`C68_185_05_meanbin`,`C68_185_06_meanbin`),
                            x.vals=samples_C68_185$conc,
                            count.vals=c(`C68_185_01_totalcount`,`C68_185_02_totalcount`,`C68_185_03_totalcount`,
                                         `C68_185_04_totalcount`,`C68_185_05_totalcount`,`C68_185_06_totalcount`)),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_3 data for each barcode
dt[,c("EC50_C68_3","EC50_SE_C68_3","response_C68_3","baseline_C68_3","nMSR_C68_3") :=
     tryCatch(fit.titration(y.vals=c(`C68_3_01_meanbin`,`C68_3_02_meanbin`,`C68_3_03_meanbin`,
                                     `C68_3_04_meanbin`,`C68_3_05_meanbin`,`C68_3_06_meanbin`),
                            x.vals=samples_C68_3$conc,
                            count.vals=c(`C68_3_01_totalcount`,`C68_3_02_totalcount`,`C68_3_03_totalcount`,
                                         `C68_3_04_totalcount`,`C68_3_05_totalcount`,`C68_3_06_totalcount`)),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_10 data for each barcode
dt[,c("EC50_C68_10","EC50_SE_C68_10","response_C68_10","baseline_C68_10","nMSR_C68_10") :=
     tryCatch(fit.titration(y.vals=c(`C68_10_01_meanbin`,`C68_10_02_meanbin`,`C68_10_03_meanbin`,
                                     `C68_10_04_meanbin`,`C68_10_05_meanbin`,`C68_10_06_meanbin`),
                            x.vals=samples_C68_10$conc,
                            count.vals=c(`C68_10_01_totalcount`,`C68_10_02_totalcount`,`C68_10_03_totalcount`,
                                         `C68_10_04_totalcount`,`C68_10_05_totalcount`,`C68_10_06_totalcount`),
                            EC50.start=1),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_88 data for each barcode
dt[,c("EC50_C68_88","EC50_SE_C68_88","response_C68_88","baseline_C68_88","nMSR_C68_88") :=
     tryCatch(fit.titration(y.vals=c(`C68_88_01_meanbin`,`C68_88_02_meanbin`,`C68_88_03_meanbin`,
                                     `C68_88_04_meanbin`,`C68_88_05_meanbin`,`C68_88_06_meanbin`),
                            x.vals=samples_C68_88$conc,
                            count.vals=c(`C68_88_01_totalcount`,`C68_88_02_totalcount`,`C68_88_03_totalcount`,
                                         `C68_88_04_totalcount`,`C68_88_05_totalcount`,`C68_88_06_totalcount`)),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_201 data for each barcode
dt[,c("EC50_C68_201","EC50_SE_C68_201","response_C68_201","baseline_C68_201","nMSR_C68_201") :=
     tryCatch(fit.titration(y.vals=c(`C68_201_01_meanbin`,`C68_201_02_meanbin`,`C68_201_03_meanbin`,
                                     `C68_201_04_meanbin`,`C68_201_05_meanbin`,`C68_201_06_meanbin`),
                            x.vals=samples_C68_201$conc,
                            count.vals=c(`C68_201_01_totalcount`,`C68_201_02_totalcount`,`C68_201_03_totalcount`,
                                         `C68_201_04_totalcount`,`C68_201_05_totalcount`,`C68_201_06_totalcount`),
                            EC50.start=1),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_239 data for each barcode
dt[,c("EC50_C68_239","EC50_SE_C68_239","response_C68_239","baseline_C68_239","nMSR_C68_239") :=
     tryCatch(fit.titration(y.vals=c(`C68_239_01_meanbin`,`C68_239_02_meanbin`,`C68_239_03_meanbin`,
                                     `C68_239_04_meanbin`,`C68_239_05_meanbin`,`C68_239_06_meanbin`),
                            x.vals=samples_C68_239$conc,
                            count.vals=c(`C68_239_01_totalcount`,`C68_239_02_totalcount`,`C68_239_03_totalcount`,
                                         `C68_239_04_totalcount`,`C68_239_05_totalcount`,`C68_239_06_totalcount`)),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]


#fit titration to C68_490 data for each barcode
dt[,c("EC50_C68_490","EC50_SE_C68_490","response_C68_490","baseline_C68_490","nMSR_C68_490") :=
     tryCatch(fit.titration(y.vals=c(`C68_490_01_meanbin`,`C68_490_02_meanbin`,`C68_490_03_meanbin`,
                                     `C68_490_04_meanbin`,`C68_490_05_meanbin`,`C68_490_06_meanbin`),
                            x.vals=samples_C68_490$conc,
                            count.vals=c(`C68_490_01_totalcount`,`C68_490_02_totalcount`,`C68_490_03_totalcount`,
                                         `C68_490_04_totalcount`,`C68_490_05_totalcount`,`C68_490_06_totalcount`),
                            EC50.start=1),
              error=function(e){list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))}),by=c("library","barcode")]

```
## QC and sanity checks
We will do some QC to make sure we got good titration curves for most of our library barcodes. We will also spot check titration curves from across our measurement range, and spot check curves whose fit parameters hit the different boundary conditions of the fit variables.

We successfully generated EC50 estimates for `r sum(!is.na(dt$EC50_C68_61))/length(!is.na(dt$EC50_C68_61))` of our C68_61 mAb measurements, `r sum(!is.na(dt$EC50_C68_183))/length(!is.na(dt$EC50_C68_183))` of the C68_183 mAb, `r sum(!is.na(dt$EC50_C68_185))/length(!is.na(dt$EC50_C68_185))` of C68_185, `r sum(!is.na(dt$EC50_C68_3))/length(!is.na(dt$EC50_C68_3))` of C68_3 mAb, `r sum(!is.na(dt$EC50_C68_10))/length(!is.na(dt$EC50_C68_10))` of our C68_10 mAb, `r sum(!is.na(dt$EC50_C68_88))/length(!is.na(dt$EC50_C68_88))` of our C68_88 mAb, `r sum(!is.na(dt$EC50_C68_201))/length(!is.na(dt$EC50_C68_201))` of our C68_201 mAb, `r sum(!is.na(dt$EC50_C68_239))/length(!is.na(dt$EC50_C68_239))` of our C68_239 mAb, and `r sum(!is.na(dt$EC50_C68_490))/length(!is.na(dt$EC50_C68_490))` of our C68_490 mAb.


```{r plot_titration_functions}
#make functions that allow me to plot a titration for any given row from the counts data frames, for spot checking curves
plot.titration <- function(row,mAb,output.text=F){
  y.vals <- c();for(sample in get(paste("samples_",mAb,sep=""))$sample){y.vals <- c(y.vals,paste(sample,"_meanbin",sep=""))};y.vals <- unlist(dt[row,y.vals,with=F])
  x.vals <- get(paste("samples_",mAb,sep=""))$conc
  count.vals <- c();for(sample in get(paste("samples_",mAb,sep=""))$sample){count.vals <- c(count.vals,paste(sample,"_totalcount",sep=""))};count.vals <- unlist(dt[row,count.vals,with=F])
  title <- paste(dt[row,target]," ",mAb,sep="")
  indices <- count.vals>cutoff & !is.na(count.vals)
  y.vals <- y.vals[indices]
  x.vals <- x.vals[indices]
  plot(x.vals,y.vals,xlab=paste("[",mAb,"] (ng/mL)",sep=""),
       ylab="mean bin",log="x",ylim=c(1,4),xlim=c(0.01,11000),pch=19,main=title)
  EC50_var <- paste("EC50_",mAb,sep="")
  fit <- nls(y.vals ~ (a/(1+(EC50/x.vals)^n))+b,
             start=list(a=3,b=1,EC50=dt[row,get(EC50_var)],n=1),
             lower=list(a=2,b=1,EC50=0.001,n=1/2),
             upper=list(a=3,b=1.25,EC50=100000,n=2),
             algorithm="port")
  if(!is.na(dt[row,get(EC50_var)])){
    lines(10^c(seq(-3,5,0.25)),predict(fit,newdata=list(x.vals=10^c(seq(-3,5,0.25)))))
    legend("topleft",bty="n",cex=1,legend=paste("EC50",format(dt[row,get(EC50_var)],digits=3),"ng/mL"))
  }
  if(output.text==T){ #for troubleshooting and interactive work, output some info from the counts table for the given row
    vars <- c("barcode","variant_class","wildtype","position","mutant",as.character(paste(mAb,"_avgcount",sep="")),as.character(paste(mAb,"_min_cell_filtered",sep="")),as.character(paste("EC50_",mAb,sep="")),as.character(paste("EC50_SE_",mAb,sep="")),as.character(paste("baseline_",mAb,sep="")),as.character(paste("response_",mAb,sep="")),as.character(paste("nMSR_",mAb,sep="")))
    return(dt[row,..vars])
  }
}
```

```{r EC50_C68_61, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_61>=10000)[1],"C68_61")
plot.titration(which(dt$EC50_C68_61>=10000)[2],"C68_61")
plot.titration(which(dt$EC50_C68_61>=10000)[3],"C68_61")
plot.titration(which(dt$EC50_C68_61>1 & dt$EC50_C68_61<10)[1],"C68_61")
plot.titration(which(dt$EC50_C68_61>1 & dt$EC50_C68_61<10)[2],"C68_61")
plot.titration(which(dt$EC50_C68_61>1 & dt$EC50_C68_61<10)[3],"C68_61")
plot.titration(which(dt$EC50_C68_61<0.5)[1],"C68_61")
plot.titration(which(dt$EC50_C68_61<0.5)[2],"C68_61")
plot.titration(which(dt$EC50_C68_61<0.5)[3],"C68_61")

```


```{r EC50_C68_183, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_183>=10000)[1],"C68_183")
plot.titration(which(dt$EC50_C68_183>=10000)[2],"C68_183")
plot.titration(which(dt$EC50_C68_183>=10000)[3],"C68_183")
plot.titration(which(dt$EC50_C68_183>1 & dt$EC50_C68_183<10)[1],"C68_183")
plot.titration(which(dt$EC50_C68_183>1 & dt$EC50_C68_183<10)[2],"C68_183")
plot.titration(which(dt$EC50_C68_183>1 & dt$EC50_C68_183<10)[3],"C68_183")
plot.titration(which(dt$EC50_C68_183<1)[1],"C68_183")
plot.titration(which(dt$EC50_C68_183<1)[2],"C68_183")
plot.titration(which(dt$EC50_C68_183<1)[3],"C68_183")

```
```{r EC50_C68_185, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_185>=10000)[1],"C68_185")
plot.titration(which(dt$EC50_C68_185>=10000)[2],"C68_185")
plot.titration(which(dt$EC50_C68_185>=10000)[3],"C68_185")
plot.titration(which(dt$EC50_C68_185>1 & dt$EC50_C68_185<10)[1],"C68_185")
plot.titration(which(dt$EC50_C68_185>1 & dt$EC50_C68_185<10)[2],"C68_185")
plot.titration(which(dt$EC50_C68_185>1 & dt$EC50_C68_185<10)[3],"C68_185")
plot.titration(which(dt$EC50_C68_185<0.5)[1],"C68_185")
plot.titration(which(dt$EC50_C68_185<0.5)[2],"C68_185")
plot.titration(which(dt$EC50_C68_185<0.5)[3],"C68_185")

```

```{r EC50_C68_3, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_3>=10000)[1],"C68_3")
plot.titration(which(dt$EC50_C68_3>=10000)[2],"C68_3")
plot.titration(which(dt$EC50_C68_3>=10000)[3],"C68_3")
plot.titration(which(dt$EC50_C68_3>1 & dt$EC50_C68_3<10)[1],"C68_3")
plot.titration(which(dt$EC50_C68_3>1 & dt$EC50_C68_3<10)[2],"C68_3")
plot.titration(which(dt$EC50_C68_3>1 & dt$EC50_C68_3<10)[3],"C68_3")
plot.titration(which(dt$EC50_C68_3<0.5)[1],"C68_3")
plot.titration(which(dt$EC50_C68_3<0.5)[2],"C68_3")
plot.titration(which(dt$EC50_C68_3<0.5)[3],"C68_3")

```

```{r EC50_C68_10, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_10>=10000)[1],"C68_10")
plot.titration(which(dt$EC50_C68_10>=10000)[2],"C68_10")
plot.titration(which(dt$EC50_C68_10>=10000)[3],"C68_10")
plot.titration(which(dt$EC50_C68_10>1 & dt$EC50_C68_10<10)[1],"C68_10")
plot.titration(which(dt$EC50_C68_10>1 & dt$EC50_C68_10<10)[2],"C68_10")
plot.titration(which(dt$EC50_C68_10>1 & dt$EC50_C68_10<10)[3],"C68_10")
plot.titration(which(dt$EC50_C68_10<0.5)[1],"C68_10")
plot.titration(which(dt$EC50_C68_10<0.5)[2],"C68_10")
plot.titration(which(dt$EC50_C68_10<0.5)[3],"C68_10")

```

```{r EC50_C68_88, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_88>=10000)[1],"C68_88")
plot.titration(which(dt$EC50_C68_88>=10000)[2],"C68_88")
plot.titration(which(dt$EC50_C68_88>=10000)[3],"C68_88")
plot.titration(which(dt$EC50_C68_88>1 & dt$EC50_C68_88<10)[1],"C68_88")
plot.titration(which(dt$EC50_C68_88>1 & dt$EC50_C68_88<10)[2],"C68_88")
plot.titration(which(dt$EC50_C68_88>1 & dt$EC50_C68_88<10)[3],"C68_88")
plot.titration(which(dt$EC50_C68_88<0.5)[1],"C68_88")
plot.titration(which(dt$EC50_C68_88<0.5)[2],"C68_88")
plot.titration(which(dt$EC50_C68_88<0.5)[3],"C68_88")

```

```{r EC50_C68_201, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_201>=10000)[1],"C68_201")
plot.titration(which(dt$EC50_C68_201>=10000)[2],"C68_201")
plot.titration(which(dt$EC50_C68_201>=10000)[3],"C68_201")
plot.titration(which(dt$EC50_C68_201>1 & dt$EC50_C68_201<10)[1],"C68_201")
plot.titration(which(dt$EC50_C68_201>1 & dt$EC50_C68_201<10)[2],"C68_201")
plot.titration(which(dt$EC50_C68_201>1 & dt$EC50_C68_201<10)[3],"C68_201")
plot.titration(which(dt$EC50_C68_201<0.5)[1],"C68_201")
plot.titration(which(dt$EC50_C68_201<0.5)[2],"C68_201")
plot.titration(which(dt$EC50_C68_201<0.5)[3],"C68_201")

```

```{r EC50_C68_239, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_239>=10000)[1],"C68_239")
plot.titration(which(dt$EC50_C68_239>=10000)[2],"C68_239")
plot.titration(which(dt$EC50_C68_239>=10000)[3],"C68_239")
plot.titration(which(dt$EC50_C68_239>1 & dt$EC50_C68_239<10)[1],"C68_239")
plot.titration(which(dt$EC50_C68_239>1 & dt$EC50_C68_239<10)[2],"C68_239")
plot.titration(which(dt$EC50_C68_239>1 & dt$EC50_C68_239<10)[3],"C68_239")
plot.titration(which(dt$EC50_C68_239<0.5)[1],"C68_239")
plot.titration(which(dt$EC50_C68_239<0.5)[2],"C68_239")
plot.titration(which(dt$EC50_C68_239<0.5)[3],"C68_239")

```

```{r EC50_C68_490, fig.width=8, fig.height=8, fig.align="center",dpi=300, warning=FALSE, results=FALSE,dev="png"}
par(mfrow=c(3,3))
plot.titration(which(dt$EC50_C68_490>=10)[1],"C68_490")
plot.titration(which(dt$EC50_C68_490>=10)[2],"C68_490")
plot.titration(which(dt$EC50_C68_490>=10)[3],"C68_490")
plot.titration(which(dt$EC50_C68_490>1 & dt$EC50_C68_490<10)[1],"C68_490")
plot.titration(which(dt$EC50_C68_490>1 & dt$EC50_C68_490<10)[2],"C68_490")
plot.titration(which(dt$EC50_C68_490>1 & dt$EC50_C68_490<10)[3],"C68_490")
plot.titration(which(dt$EC50_C68_490<0.5)[1],"C68_490")
plot.titration(which(dt$EC50_C68_490<0.5)[2],"C68_490")
plot.titration(which(dt$EC50_C68_490<0.5)[3],"C68_490")

```

## Data filtering by fit quality

Next, let's filter out poor fits using the value we previously computed, the *normalized* mean square residual (nMSR). This metric computes the residual between the observed response variable and that predicted from the titration fit, normalizes this residual by the response range of the titration fit (which is allowed to vary between 1.5 and 3 per the titration fits above), and computes the mean-square of these normalized residuals.

Distribution of the nMSR metric in each set of fits

```{r nMSR_distribution, fig.width=9, fig.height=9, fig.align="center",dpi=300,dev="png"}
par(mfrow=c(3,3))
hist(dt$nMSR_C68_61,main="C68_61",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_183,main="C68_183",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_185,main="C68_185",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_3,main="C68_3",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_10,main="C68_10",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_88,main="C68_88",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_201,main="C68_201",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_239,main="C68_239",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
hist(dt$nMSR_C68_490,main="C68_490",xlab="Response-normalized mean squared residual",col="gray50",breaks=40,xlim=c(0,0.6))
```

As we would expect, the MSR stat decreases with cell count, indicating that higher cell counts leads to better curve fits. Also show the cutoff I'm proposing for nMSR (10x median across all fits), legend gives percent of curve fits eliminated
```{r nMSR_v_cell_count, fig.width=13.5, fig.height=18, fig.align="center",dpi=300,dev="png"}
median.nMSR <- median(c(dt$nMSR_C68_61,dt$nMSR_C68_183,dt$nMSR_C68_185,dt$nMSR_C68_3),na.rm=T)

par(mfrow=c(3,3))
plot(log10(dt$`C68_61_avgcount`),dt$nMSR_C68_61,main="C68_61",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_61 > 10*median.nMSR & !is.na(nMSR_C68_61),])/nrow(dt[!is.na(nMSR_C68_61),]),digits=3),"%"))

plot(log10(dt$`C68_183_avgcount`),dt$nMSR_C68_183,main="C68_183",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_183 > 10*median.nMSR & !is.na(nMSR_C68_183),])/nrow(dt[!is.na(nMSR_C68_183),]),digits=3),"%"))

plot(log10(dt$`C68_185_avgcount`),dt$nMSR_C68_185,main="C68_185",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_185 > 10*median.nMSR & !is.na(nMSR_C68_185),])/nrow(dt[!is.na(nMSR_C68_185),]),digits=3),"%"))

plot(log10(dt$`C68_3_avgcount`),dt$nMSR_C68_3,main="C68_3",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_3 > 10*median.nMSR & !is.na(nMSR_C68_3),])/nrow(dt[!is.na(nMSR_C68_3),]),digits=3),"%"))

plot(log10(dt$`C68_10_avgcount`),dt$nMSR_C68_10,main="C68_10",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_10 > 10*median.nMSR & !is.na(nMSR_C68_10),])/nrow(dt[!is.na(nMSR_C68_10),]),digits=3),"%"))

plot(log10(dt$`C68_88_avgcount`),dt$nMSR_C68_88,main="C68_88",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_88 > 10*median.nMSR & !is.na(nMSR_C68_88),])/nrow(dt[!is.na(nMSR_C68_88),]),digits=3),"%"))

plot(log10(dt$`C68_201_avgcount`),dt$nMSR_C68_201,main="C68_201",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_201 > 10*median.nMSR & !is.na(nMSR_C68_201),])/nrow(dt[!is.na(nMSR_C68_201),]),digits=3),"%"))

plot(log10(dt$`C68_239_avgcount`),dt$nMSR_C68_239,main="C68_239",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_239 > 10*median.nMSR & !is.na(nMSR_C68_239),])/nrow(dt[!is.na(nMSR_C68_239),]),digits=3),"%"))

plot(log10(dt$`C68_490_avgcount`),dt$nMSR_C68_490,main="C68_490",pch=19,col="#00000010",xlab="average cell count (log10)",ylab="nMSR",xlim=c(1,3),ylim=c(0,0.6))
abline(h=10*median.nMSR,col="red",lty=2)
legend("topleft",bty="n",cex=1,legend=paste(format(100*nrow(dt[nMSR_C68_490 > 10*median.nMSR & !is.na(nMSR_C68_490),])/nrow(dt[!is.na(nMSR_C68_490),]),digits=3),"%"))
```


Next, we will apply this filtering step on normalized MSR, removing curves with nMSR >10x the median across all experiments
```{r filter_curves_by_nMSR}
dt[nMSR_C68_61 > 10*median.nMSR,c("EC50_C68_61","EC50_SE_C68_61","response_C68_61","baseline_C68_61") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_183 > 10*median.nMSR,c("EC50_C68_183","EC50_SE_C68_183","response_C68_183","baseline_C68_183") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_185 > 10*median.nMSR,c("EC50_C68_185","EC50_SE_C68_185","response_C68_185","baseline_C68_185") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_3 > 10*median.nMSR,c("EC50_C68_3","EC50_SE_C68_3","response_C68_3","baseline_C68_3") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_10 > 10*median.nMSR,c("EC50_C68_10","EC50_SE_C68_10","response_C68_10","baseline_C68_10") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_88 > 10*median.nMSR,c("EC50_C68_88","EC50_SE_C68_88","response_C68_88","baseline_C68_88") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_201 > 10*median.nMSR,c("EC50_C68_201","EC50_SE_C68_201","response_C68_201","baseline_C68_201") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_239 > 10*median.nMSR,c("EC50_C68_239","EC50_SE_C68_239","response_C68_239","baseline_C68_239") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

dt[nMSR_C68_490 > 10*median.nMSR,c("EC50_C68_490","EC50_SE_C68_490","response_C68_490","baseline_C68_490") := list(as.numeric(NA),as.numeric(NA),as.numeric(NA),as.numeric(NA))]

```

## Final scores

Let's visualize the EC50 binding measurements as violin plots for the different wildtype targets, for each mAb.

```{r binding_distribution_vioplot_C68_61, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_61),],aes(x=target,y=EC50_C68_61))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_61 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_61.pdf",sep="")))
```


```{r binding_distribution_vioplot_C68_183, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_183),],aes(x=target,y=EC50_C68_183))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_183 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_183.pdf",sep="")))
```


```{r binding_distribution_vioplot_C68_185, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_185),],aes(x=target,y=EC50_C68_185))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_185 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_185.pdf",sep="")))
```



```{r binding_distribution_vioplot_C68_3, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_3),],aes(x=target,y=EC50_C68_3))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_3 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_3.pdf",sep="")))
```


```{r binding_distribution_vioplot_C68_10, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_10),],aes(x=target,y=EC50_C68_10))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_10 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_10.pdf",sep="")))
```

```{r binding_distribution_vioplot_C68_88, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_88),],aes(x=target,y=EC50_C68_88))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_88 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_88.pdf",sep="")))
```

```{r binding_distribution_vioplot_C68_201, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_201),],aes(x=target,y=EC50_C68_201))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_201 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_201.pdf",sep="")))
```

```{r binding_distribution_vioplot_C68_239, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_239),],aes(x=target,y=EC50_C68_239))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_239 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_239.pdf",sep="")))
```

```{r binding_distribution_vioplot_C68_490, echo=T, fig.width=15, fig.height=5, fig.align="center", dpi=300,dev="png"}
p1 <- ggplot(dt[!is.na(EC50_C68_490),],aes(x=target,y=EC50_C68_490))+
  geom_violin(scale="width")+stat_summary(fun=median,geom="point",size=1)+
  ggtitle("C68_490 EC50")+xlab("target")+theme(axis.text.x=element_text(angle=-90,hjust=0))+
  facet_wrap(~library,nrow=1)+
  scale_y_log10()

grid.arrange(p1,ncol=1)

#save pdf
invisible(dev.print(pdf, paste(config$mAb_EC50_dir,"/violin-plot_EC50-by-target_C68_490.pdf",sep="")))
```
## Save barcode-level metrics

In the next script, we will collapse bcs down to final mutant/variant-level phenotypes, integrate things like expression effects of variants, and visualize final phenotypes.

```{r output_data_barcode-level}
dt[,.(library,barcode,target,variant_class,
     `C68_61_avgcount`,EC50_C68_61,
     `C68_183_avgcount`,EC50_C68_183,
     `C68_185_avgcount`,EC50_C68_185,
     `C68_3_avgcount`,EC50_C68_3,
     `C68_10_avgcount`,EC50_C68_10,
     `C68_88_avgcount`,EC50_C68_88,
     `C68_201_avgcount`,EC50_C68_201,
     `C68_239_avgcount`,EC50_C68_239,
     `C68_490_avgcount`,EC50_C68_490)] %>%
  mutate_if(is.numeric, round, digits=6) %>%
  write.csv(file=config$mAb_EC50_file, row.names=F)

```



